# "Reproducible Research: Peer Assessment 1"

```{r LoadPackages,echo=FALSE,message=FALSE}
library(dplyr)
library(lattice)
```


## Loading and preprocessing the data
```{r LoadData}
activity <- read.csv("activity.csv", stringsAsFactors=FALSE)
activity$date <- as.Date(activity$date,"%Y-%m-%d")
head(activity)
```



## What is mean total number of steps taken per day?
```{r HistSteps, fig.width=7}
stepsperday <- by(activity$steps,activity$date,sum)
meansteps <- mean(stepsperday,na.rm=TRUE)
mediansteps <- median(stepsperday,na.rm=TRUE)
hist(stepsperday,breaks=30,xlab="Number of steps per day",
     main = "Histogram of total steps per day", col=rgb(0.094,0.455,0.804))
abline(v=meansteps,lwd=2)
text(meansteps,10,pos=4,"Mean")
```

The mean total number of steps per day is **`r as.integer(meansteps)`** and the median total number
of steps per day is **`r mediansteps`**.


## What is the average daily activity pattern?
```{r StepsPerInterval, fig.width=9}
meansteps = by(activity$steps,activity$interval,mean,na.rm=TRUE)
highint <- names(which.max(meansteps))
plot(names(meansteps),meansteps,type="l",lwd=2,xlab="Interval",ylab="Steps",main="Average daily steps per interval")
abline(v=as.numeric(highint),col=rgb(0.094,0.455,0.804),lwd=2)
text(as.numeric(highint),200,pos=4,"Interval with the highest number of steps",col=rgb(0.094,0.455,0.804))
```

Interval number **`r highint`** has the highest average daily number of steps (`r as.integer(meansteps[highint])`).

## Imputing missing values
I have chosen to impute all NAs with the mean for the corresponding interval. 

```{r HistStepsNoNAs,fig.width=7}
noNA <- sum(is.na(activity$steps))
newActivity <- activity %>% 
    group_by(interval) %>% 
    mutate(steps = ifelse(is.na(steps), as.integer(round(mean(steps,na.rm=TRUE))), steps))
stepsperday <- by(newActivity$steps,newActivity$date,sum)
meanperday = mean(stepsperday,na.rm=TRUE)
medianperday = median(stepsperday,na.rm=TRUE)
hist(stepsperday,breaks=30,xlab="Number of steps per day",main = "Histogram of total steps per day",
     col=rgb(0.094,0.455,0.804))
abline(v=meanperday,lwd=2)
text(meanperday,15,pos=4,"Mean")
```

The new mean total number of steps per day is `r as.integer(meanperday)` and the median total number
of steps per day is `r medianperday`. These values do not differ significantly from the mean and median calculated before. Imputing missing data with the mean for that interval did only increase the frequency of mean/median level total number of steps per day. 


## Are there differences in activity patterns between weekdays and weekends?
```{r DayOfTheWeek, fig.width=9}
newActivity$dayofweek <- factor(ifelse(weekdays(newActivity$date) %in% 
                                           c("Saturday","Sunday"),"Weekend","Weekday"))
meanActivity <- newActivity %>% group_by(dayofweek,interval) %>% summarise(steps = mean(steps))
xyplot(steps~interval|dayofweek,data = meanActivity,type="l",lwd=2,col="black", xlab="Interval",
       ylab="Number of Steps", par.settings=list(strip.background=list(col=rgb(0.094,0.455,0.804,0.8))),
       layout=c(1,2))
```


